---
title: "DATA 101 Exam 2"
author: "Kamila Palys"
date: "Due: Sunday, 11/29 at 11:59pm"
output: pdf_document
---

## Academic Honesty Statement (fill in your name)

I, Kamila Palys, hereby affirm that I have not communicated with or gained information in any way from my classmates or anyone other than the Professor during this exam, that I have not assisted anyone else with this exam, and that all work is my own.

## Load packages and data

```{r load-packages, message=FALSE}
# load required packages here
library(tidyverse)
library(tidymodels)
library(janitor)
library(lubridate)
library(NHANES)
```

```{r read-data, message=FALSE}
# read in the data here
nhanes <- NHANES %>%
  janitor::clean_names()
```

## Questions 

### Question 1

Here we will show the relationship between the height and body mass index of a person using a scatterplot. 

```{r bmi vs height}
ggplot(data = nhanes, mapping = aes(x = height, y = bmi)) +
  geom_point(alpha = 0.2) + 
  labs(x = "Height in cm", 
       y = "Body Mass Index (BMI)", 
       title = "Relationship Between Height and Body Mass Index") +
  theme_minimal()
```

From this display, a slight upwards trend is seen. For heights of less than 150cm, not much of a slope is visible and body mass index stays relatively the same for various heights under 150cm. For heights above that amount, there are many more people whose body mass indexes are higher and there is a far greater range of them. In fact, most of the body mass indexes are higher than for people shorter than 150cm. One reason as to why this may be the case is because those shorter than 150cm are usually young children, whose parents have the main say in their diets. When the children grow up and therefore also get taller, they make more of their own decisions when it comes to their diet, which may not be good for their health and BMI. 

Now we will show the relationship between the weight and body mass index of a person, also using a scatterplot. 

```{r bmi vs weight}
ggplot(data = nhanes, mapping = aes(x = weight, y = bmi)) +
  geom_point(alpha = 0.15) + 
  labs(x = "Weight (kg)", 
       y = "Body Mass Index (BMI)", 
       title = "Relationship Between Weight and Body Mass Index") + 
  theme_minimal()
```

Here, there is almost a perfectly linear positive slope between a person's weight and their body mass index. The positive relationship makes sense because as the weight of a person goes up, given a height, their body mass index will go up as well. Even though height is not a factor shown on this graph, it often increases along with weight, which may lead to little to no change in body mass index. However, weight can fluctuate a lot more and throughout a person's entire life, which is why this upward trend is seen, suggesting that a higher weight still generally leads to a higher body mass index. 

### Question 2

Now, a linear model will be created to predict bmi as a function of weight.  

```{r linear model}
bmi_fit <- lm(bmi ~ weight, data = nhanes)
tidy(bmi_fit)
```

$\hat{bmi} = 9.09 + 0.24weight$

The coefficient of weight in this equation is 0.24, rounded to the hundredths, meaning that for every additional kilogram of weight, a person's bmi is expected to go up by approximately 0.24. 

### Question 3

Now the coefficient of determination, or R-Squared, will be calculated.

```{r}
glance(bmi_fit)$r.squared
```

The coefficient of determination being approximately 0.81 tells us that 81% of the variability of the bmi can be explained by the weight from the linear model previously created. This high percentage suggests that the model created is a good fit for the data. 

### Question 4

[Enter code and/or narrative here.]

### Question 5

[Enter code and/or narrative here.]

### Question 6

[Enter code and/or narrative here.]

### Question 7

[Enter code and/or narrative here.]

### Question 8

[Enter code and/or narrative here.]

### Question 9

[Enter code and/or narrative here.]

### Question 10

[Enter code and/or narrative here.]

### Question 11

[Enter code and/or narrative here.]

### Question 12

[Enter code and/or narrative here.]

### Question 13

[Enter code and/or narrative here.]
